exp1_lik_by_trial

### 2020-09-08 Experiment_2/exp2_hmm_strategies.Rmd
fixed bugs in report

```{r}
# we need to have "dat" available and lik_by_trial
QLS_lik <- lik_by_trial$QLS_lik
QLS_lik[which(QLS_lik == 0)] <- .001
QLS_lik[is.na(QLS_lik)] <- .33 # this is a hack; there shouldn't be any missing values
Bayes_NT_lik <- lik_by_trial$Bayes_NT_lik
Bayes_Tr_lik <- lik_by_trial$Bayes_Tr_lik

nsubject <- length(unique(dat$human_id)) # number of participants
ngame <- 3 # number of games
ntrial <- c(50,50,50) # numer of trials in each game

@@

fmod <- fit(mod, emcontrol=em.control(random.start=FALSE))

@@

>>>>

```{r}
# we need to have "dat" available and lik_by_trial
QLS_lik <- lik_by_trial_2$QLS_lik
Bayes_NT_lik <- lik_by_trial_2$Bayes_NT_lik
Bayes_Tr_lik <- lik_by_trial_2$Bayes_Tr_lik

nsubject <- length(unique(lik_by_trial_2$human_id)) # number of participants
ngame <- 12 # number of games
#ntrial <- c(60,60,60)

ntrial <- c(20,20,10,10,20,20,10,10,20,20,10,10) # number of trials in each game

@@

fmod <- fit(mod, emcontrol=em.control(random.start=TRUE))

summary(fmod,compact=FALSE)
```

```{r}

# Create table for each trial and calc proportions of states 
# Check book draft for weather prediction data 
# p <- posterior(fmod)[60:66,1]
# p

forwardbackward(fmod)$gamma
#check section called decoding

```{r}

@@

# Use bootstrapped likeihood ratio test....(see chapter 2)

### 2020-09-08 Report/draft_report_v2.Rmd
fixed bugs in report

---
title             : "Transfer of Learned Opponent Models in Zero Sum Games"
shorttitle        : "Draft-1"

@@

```{r load-packages, include = FALSE}
library(papaja)
require(knitr)
require(citr)
require(bookdown)
# using some functions dplyr, ggpubr, PairedData and sjPlot. Need to be loaded. 
library(tidyr)
library(dplyr)
library(MASS)
library(ggpubr)
library(afex)
library(PairedData)
library(multcompView)
library(lsmeans)
library(magick)
```

@@

Finally, the analysis showed a significant interaction effects of block by game ( F(2,100) = 6.92 , p = .002, $\eta^{2}$ = 0.02), indicating that within-game learning differed between the games. Indeed, second half scores in RPS are significantly higher than first-half scores (t(150) = 5.59, p <.0001), while there was no significant difference between block scores for the other two games. This is indicative of the significant within game transfer within the first game when players have no epxerience against the opponent, as opposed to much lower within game performance improvement in the latter games when participants have had some exprience playing against the opponenet and start with higher scores indicative of transfer.  There was also a three-way interaction between condition, game, and block ( F(2,100) = 3.88 , p = .023, $\eta^{2}$ = 0.01), which indicates, as seen in Figure\ \@ref(fig:exp1-3factor-plot) that within-game learning changes across games also depend on the sophistication of the opponent. For instance, there is more within game learning in the third game against level-2 opponents, since the initial scores are lower than agaisnt level-1 opponent. The explanation for this will become clearer when we discuss the factors moderating learning transfer in the next section.

@@

As a measure for learning transfer, we focus on participants’ scores in rounds 2-6, excluding the very first round for which the agent is programmed to play randomly as it has no data on prior rounds on which to build its response. A group of players with no experience of the game are expected to have scores not significantly different from 0. Any significantly positive group average scores would therefore reflect prior learning from past experiences.

@@

apa_lm1 <- apa_print(exp1_early_anova)

@@

All_comp_results <- read.csv(file="../All_results.csv")
exp1_comp_table <- table(All_comp_results[, "condition"],c("Random","ToM_BT","ToM_NBT","QL", "QL_states","EWA","S_EWA","EWA_states", "MBM")[apply(All_comp_results[,c("Random_BIC","Bayes_Tr_BIC","Bayes_No_Tr_BIC","QL_BIC","QL_states_BIC","EWA_BIC","EWA_self_BIC","EWA_states_BIC")],1,which.min)])

@@

exp1_model_comp <- data.frame()
for(id in unique(exp1_dat$human_id)) {
  tdat <- subset(exp1_dat,human_id == id)
  tot_score <- sum(tdat$score)
  tot_time <- sum(tdat$human_rt)
  early_dat <- subset(tdat,between(tdat$round,2,6) & (game =="fwg"))
  #early_dat <- subset(tdat,between(tdat$round,2,6) & (game =="fwg" | game =="numbers") )
  tr_score <- sum(early_dat$score)
  id_results <- subset(All_comp_results, ID == id)
  min_BIC <- apply(id_results[,c("Random_BIC","Bayes_Tr_BIC","Bayes_No_Tr_BIC","QL_BIC","QL_states_BIC","EWA_BIC","EWA_self_BIC")],1,min)

@@

```{r exp1-comp-models, fig.cap = "Histogram of best fitting computational models by condition for experiment 1", fig.align='center'}

@@

```{r exp1-cumScores, fig.cap = "Average cumulative scores of participants in experiment 1 by best fitting model", fig.align='center'}
p5 <- ggplot(data = dat_by_model, aes(x = t, y=model_acc_sum, group = Best_model)) + 
   geom_line(aes(color= Best_model))

p5+ labs(color = "Best fitting model", x = "Round Number", y="Accumulated score")
```

@@

```{r exp2-comp-models, fig.cap = "Histogram of best fitting computational models by condition for experiment 2", fig.align='center'}

#knitr::include_graphics("../Report/images/exp1_comp_models.png", dpi = 108)
par(las=2) # make label text perpendicular to axis
par(mar=c(5,8,4,2)) # increase y-axis margin.
barplot(exp2_comp_table,
        horiz =TRUE, # rotate barplot for better visibility 
        las=1, # change orientation x axis labels 
        cex.names=0.8, # text label size
        beside =TRUE,
        xlab="Number of participants",
        # legend = c("level-1 opp first","level-2 opp first"),
        legend = paste0(rownames(exp2_comp_table) , " faced first"),
        args.legend=list( # positioning of legend box 
        x=ncol(exp2_comp_table) + 6,
        y=max(colSums(exp2_comp_table)) +1,
        bty = "n")
 )

@@

```{r exp2-cumScores, fig.cap = "Average cumulative scores of participants in experiment 2 by best fitting model", fig.align='center'}

@@

Plotting cumulative scores by best model for experiment 2, we see very similar results looking at Figure\ \@ref(fig:exp2-cumScores), in that participants whose behavior was best fit by a ToM model of learning the oppponent strategy had the highest cumulative performance. Out of these ToM models, the one in which there is within-game but no between-game transfer (NBT) had the best cumulative performance, folowed by a model in which both within and between trasnfer of opponent models is allowed (BT). The next best model from a performance perspective was a Q-learning model with states and within game transfer, followed by ToM models where either players started learning from scratch at each stage of each game (NT) and the model in which they didn't distinguish between opponenets (Naive). As expected, random play as well as self-tuning EWA were at the bottom of cumulative performance.

>>>>

---
title             : "Transfer of Learned Opponent Models in Zero Sum Games"
shorttitle        : "Draft-2"

@@

```{r load-packages, include = FALSE}
library(papaja)
require(knitr)
require(citr)
require(bookdown)
# using some functions dplyr, ggpubr, PairedData and sjPlot. Need to be loaded. 
library(tidyr)
library(dplyr)
library(MASS)
library(ggpubr)
library(afex)
library(PairedData)
library(multcompView)
library(lsmeans)
library(magick)
library(depmixS4)
```

@@

Finally, the analysis showed a significant interaction effects of block by game ( F(2,100) = 6.92 , p = .002, $\eta^{2}$ = 0.02), indicating that within-game learning differed between the games. Indeed, second half scores in RPS are significantly higher than first-half scores (t(150) = 5.59, p <.0001), while there was no significant difference between block scores for the other two games. This is indicative of the significant within game learning happening in the first game when players have no epxerience against the opponent, as opposed to much lower within game performance improvement in the latter games when participants have had some exprience playing against the opponenet and start with higher scores indicative of transfer.  There was also a three-way interaction between condition, game, and block ( F(2,100) = 3.88 , p = .023, $\eta^{2}$ = 0.01), which indicates, as seen in Figure\ \@ref(fig:exp1-3factor-plot) that within-game learning changes across games also depend on the sophistication of the opponent. For instance, there is more within game learning in the third game against level-2 opponents, since the initial scores are lower than agaisnt level-1 opponent. The explanation for this will become clearer when we discuss the factors moderating learning transfer in the next section.

@@

As a measure for learning transfer, we focus on participants’ scores in the first 5 rounds excluding the initial round (rounds 2-6). We exclude the very first round for which the agent is programmed to play randomly as it has no data on prior rounds on which to build its response, and therefore we cannot expect the participants to exhibit any learned best responses either. A group of players with no experience of the game are expected to have scores not significantly different from 0 over the early rounds of a new game, as was the case in RPS. Any significantly positive group average scores would therefore reflect prior learning from past experiences.

@@

#apa_lm1 <- apa_print(exp1_early_anova)

@@

exp1_comp_results <- read.csv(file="../exp1_all_results.csv")
exp1_comp_table <- table(exp1_comp_results[, "condition"],c("Random","ToM_BT","ToM_NBT","QL", "QL_states","EWA","S_EWA","EWA_states", "MBM")[apply(exp1_comp_results[,c("Random_BIC","Bayes_Tr_BIC","Bayes_No_Tr_BIC","QL_BIC","QL_states_BIC","EWA_BIC","EWA_self_BIC","EWA_states_BIC")],1,which.min)])

@@

exp1_model_comp <- data.frame()
for(id in unique(exp1_dat$human_id)) {
  tdat <- subset(exp1_dat,human_id == id)
  tot_score <- sum(tdat$score)
  tot_time <- sum(tdat$human_rt)
  early_dat <- subset(tdat,between(tdat$round,2,6) & (game =="fwg"))
  #early_dat <- subset(tdat,between(tdat$round,2,6) & (game =="fwg" | game =="numbers") )
  tr_score <- sum(early_dat$score)
  id_results <- subset(exp1_comp_results, ID == id)
  min_BIC <- apply(id_results[,c("Random_BIC","Bayes_Tr_BIC","Bayes_No_Tr_BIC","QL_BIC","QL_states_BIC","EWA_BIC","EWA_self_BIC")],1,min)

@@

```{r exp1-comp-models, fig.cap = "Experiment 1 - Histogram of best fitting computational models by condition", fig.align='center'}

@@

```{r exp1-cumScores, fig.cap = "Experiment 1 - Average cumulative scores of participants by best fitting model", fig.align='center'}
p5 <- ggplot(data = dat_by_model, aes(x = t, y=model_acc_sum, group = Best_model)) + 
   geom_line(aes(color= Best_model))

p5 + labs(color = "Best fitting model", x = "Round Number", y="Accumulated score")
```

However, this is at odds with the findings from the previous section regarding learning transfer: If indeed most participants use Q-learning with states to choose their actions, then they should not be able to transfer learning to the early rounds of the new game. In order to understand better what is going on, we plot the likelihood by trial for each game and each of the three strategies: Q-learning with states, and Theory of Mind models with and without the possibility of across game transfer.

Figure\ \@ref(fig:exp1-lik-by-tr) shows that in the later games, the likelihood for the ToM models is higher in the initial rounds in which learning transfer is measured, but that over time, the likelihood of Q-learning model becomes more important and exceeds that of ToM models.

```{r, include= FALSE}

exp1_lik_by_trial  <- read.csv("../exp1_lik_by_trial.csv")

mean_lik <- exp1_lik_by_trial %>%
  dplyr::select(human_id,condition,game, round, QLS_lik, Bayes_NT_lik, Bayes_Tr_lik) %>%
  group_by(condition,game,round) %>%
  summarise(mean_QLS_lik = mean(QLS_lik, na.rm = TRUE),
            mean_NT_lik = mean(Bayes_NT_lik, na.rm = TRUE),
            mean_Tr_lik = mean(Bayes_Tr_lik, na.rm = TRUE))

mean_lik$game <- factor(mean_lik$game,levels=c("rps","fwg","numbers"))

data_long <- gather(mean_lik, strategy, probability, mean_QLS_lik:mean_Tr_lik, factor_key=TRUE)

```{r exp1-lik-by-tr, fig.cap = "Experiment 1 Likelihood by trial by game and opponent faced", fig.align='center'}
# make sure the different games are ordered in the way they were played
# data_long$game <- factor(data_long$game,levels=c("rps","fwg","numbers"))
ggplot(data_long,aes(x=round,y=probability,colour=strategy)) +
  geom_line() +
  facet_grid(game~condition)

@@

```{r exp2-comp-models, fig.cap = "Experiment 2 Histogram of best fitting computational models by condition", fig.align='center'}

par(las=2) # make label text perpendicular to axis
par(mar=c(5,8,4,2)) # increase y-axis margin.
barplot(exp2_comp_table,
        horiz =TRUE, # rotate barplot for better visibility 
        las=1, # change orientation x axis labels 
        cex.names=0.8, # text label size
        beside =TRUE,
        xlab="Number of participants",
        # legend = c("level-1 opp first","level-2 opp first"),
        legend = paste0(rownames(exp2_comp_table) , " faced first"),
        args.legend=list( # positioning of legend box 
        x=ncol(exp2_comp_table) + 6,
        y=max(colSums(exp2_comp_table)) +1,
        bty = "n")
 )

@@

```{r exp2-cumScores, fig.cap = "Experiment 2 Average cumulative scores of participants by best fitting model", fig.align='center'}

@@

Plotting cumulative scores by best model for experiment 2, we see very similar results looking at Figure\ \@ref(fig:exp2-cumScores), in that participants whose behavior was best fit by a ToM model of learning the oppponent strategy had the highest cumulative performance. Out of these ToM models, the one in which there is within-game but no between-game transfer (NBT) had the best cumulative performance (although it only fit 2 participants best), followed by a model in which both within and between trasnfer of opponent models is allowed (BT). The next best model from a performance perspective was a Q-learning model with states and within game transfer, followed by ToM models where either players started learning from scratch at each stage of each game (NT) and the model in which they didn't distinguish between opponenets (Naive). As expected, random play was at the bottom of cumulative performance.

As in experiment 1, we want to understand the dynamic of strategy choice by plotting the likelihood by trial for each strategy, using the optimal parameters found when fitting the model. Figure\ \@ref(fig:exp2-lik-by-tr) shows that, as in experiment 1, ToM models had higher likelihood in the early stages of the second (most similar) game, however the likelihood of Q-learning with states models increases steadily to be the highest in the later stages of all games. In the third and mosre dissimilar game, we get a result that is different from experiment 1. In this instance, the likelihoods of the ToM models stay constant and close to their initial values.

```{r, include = FALSE }

exp2_lik_by_trial <- read.csv("../Experiment_2/exp2_lik_by_trial.csv")

mean_lik_2 <- exp2_lik_by_trial %>%
  dplyr::select(human_id,game,stage,round_condition, round, QLS_lik, QLS_lik_within, Bayes_NT_lik, Bayes_Tr_lik) %>%
  group_by(game,stage,round) %>%
  summarise(mean_QLS_lik = mean(QLS_lik, na.rm = TRUE),
            mean_QLS_within_lik = mean(QLS_lik_within, na.rm = TRUE),
            mean_NT_lik = mean(Bayes_NT_lik, na.rm = TRUE),
            mean_Tr_lik = mean(Bayes_Tr_lik, na.rm = TRUE))

mean_lik_2$game <- factor(mean_lik_2$game,levels=c("rps","fwg","shootout"))

data_long2 <- gather(mean_lik_2, strategy, probability, mean_QLS_lik:mean_Tr_lik, factor_key=TRUE)
```

```{r exp2-lik-by-tr, fig.cap = "Experiment2 likelihood by trial by game and opponent faced", fig.align ='center'}
# make sure the different games are ordered in the way they were played
# data_long2$game <- factor(data_long$game,levels=c("rps","fwg","shootout"))
ggplot(data_long2,aes(x=round,y=probability,colour=strategy)) +
  geom_line() +
  facet_grid(game~stage)

## Using Hidden Markov Model to explore strategy switching

The fact that the likelihoods of the main strategies considered cross over in both experiments could be interpreted as indicative of participants switching between strategies as the games progressed. Indeed, in both experiments, following our results, it seems that in the earlier stages of the latter games, the ToM based strategies fit observed action choices better than Q-learning based ones, with a reversal of the roles in later stages.

In order to test for the existence of strategy switching in participants' play, we fit Hidden Markov Models in which the latent states are the 3 strategies used (Q learning with state space consisting of previous round play, ToM based model with and without opponent model transfer). Hidden Markov models are useful tools to explore structure in observed time series. They are named because of two properties: First, they make the assumption that any observable action at time t results from a process whose state at time t , named $S_{t}$ is "hidden" from the observer. Second, it also assumes that this hidden process has a Markov property, meaning that given state $S_{t-1}$, the value of $S_{t}$ is independent of all states occuring before time $t-1$. We also assume that $S_{t}$ has a discrete probability distribution in that it take one of K discrete values. The model is therefore specified by initial probabilities of being in each state ${1,2,...,K}$ and transition probabilities for moving from state $i$ to state $j$. These probabilities are fit using obsserved actiosn generated from these hidden states.

Using the R DepmixS4 package, we fit a total of two HMMs. In the first model, we allow for a non-nil probability of players transitioning from one state (strategy) to another. In the second model, we assume that such switching does not happen, and as such assume implicitly that when players start with a particular strategy, they continue using it throughout the game. We then compare the likleihoods of each HMM model using a likelihood ration test. In experiment 1, the HMM model with swithcing has a significantly higher likelihood than the non-swithcing one (p-value = 0.00). The switching model in experiment two has also significantly higher likelihood (p-value = 0.00).  On top of indications from looking at the liklihood by trial graphs, we have therefore further evidence that participants did indeed switch their strategies as the games progressed.

### 2020-09-08 Report/draft_report_v2.tex
fixed bugs in report

\shorttitle{Draft-1}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Department of Experimental Psychology, University College London}
\usepackage{csquotes}
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

{\centering \includegraphics{draft_report_v1_files/figure-latex/exp1-avg-scores-game-1}

@@

{\centering \includegraphics{draft_report_v1_files/figure-latex/exp1-3factor-plot-1}

@@

Finally, the analysis showed a significant interaction effects of block by game ( F(2,100) = 6.92 , p = .002, \(\eta^{2}\) = 0.02), indicating that within-game learning differed between the games. Indeed, second half scores in RPS are significantly higher than first-half scores (t(150) = 5.59, p \textless{}.0001), while there was no significant difference between block scores for the other two games. This is indicative of the significant within game transfer within the first game when players have no epxerience against the opponent, as opposed to much lower within game performance improvement in the latter games when participants have had some exprience playing against the opponenet and start with higher scores indicative of transfer. There was also a three-way interaction between condition, game, and block ( F(2,100) = 3.88 , p = .023, \(\eta^{2}\) = 0.01), which indicates, as seen in Figure~\ref{fig:exp1-3factor-plot} that within-game learning changes across games also depend on the sophistication of the opponent. For instance, there is more within game learning in the third game against level-2 opponents, since the initial scores are lower than agaisnt level-1 opponent. The explanation for this will become clearer when we discuss the factors moderating learning transfer in the next section.

@@

As a measure for learning transfer, we focus on participants' scores in rounds 2-6, excluding the very first round for which the agent is programmed to play randomly as it has no data on prior rounds on which to build its response. A group of players with no experience of the game are expected to have scores not significantly different from 0. Any significantly positive group average scores would therefore reflect prior learning from past experiences.

@@

{\centering \includegraphics{draft_report_v1_files/figure-latex/exp1-score-by-opp-1}

@@

{\centering \includegraphics{draft_report_v1_files/figure-latex/exp2-score-by-opp-1}

@@

{\centering \includegraphics{draft_report_v1_files/figure-latex/exp2-early-scores-1}

@@

{\centering \includegraphics{draft_report_v1_files/figure-latex/exp1-comp-models-1}

@@

\caption{Histogram of best fitting computational models by condition for experiment 1}\label{fig:exp1-comp-models}
\end{figure}

@@

{\centering \includegraphics{draft_report_v1_files/figure-latex/exp1-cumScores-1}

@@

\caption{Average cumulative scores of participants in experiment 1 by best fitting model}\label{fig:exp1-cumScores}
\end{figure}

@@

{\centering \includegraphics{draft_report_v1_files/figure-latex/exp2-comp-models-1}

@@

\caption{Histogram of best fitting computational models by condition for experiment 2}\label{fig:exp2-comp-models}
\end{figure}

@@

{\centering \includegraphics{draft_report_v1_files/figure-latex/exp2-cumScores-1}

@@

\caption{Average cumulative scores of participants in experiment 2 by best fitting model}\label{fig:exp2-cumScores}
\end{figure}

Plotting cumulative scores by best model for experiment 2, we see very similar results looking at Figure~\ref{fig:exp2-cumScores}, in that participants whose behavior was best fit by a ToM model of learning the oppponent strategy had the highest cumulative performance. Out of these ToM models, the one in which there is within-game but no between-game transfer (NBT) had the best cumulative performance, folowed by a model in which both within and between trasnfer of opponent models is allowed (BT). The next best model from a performance perspective was a Q-learning model with states and within game transfer, followed by ToM models where either players started learning from scratch at each stage of each game (NT) and the model in which they didn't distinguish between opponenets (Naive). As expected, random play as well as self-tuning EWA were at the bottom of cumulative performance.

>>>>

\shorttitle{Draft-2}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Department of Experimental Psychology, University College London}
\usepackage{csquotes}
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

{\centering \includegraphics{draft_report_v2_files/figure-latex/exp1-avg-scores-game-1}

@@

{\centering \includegraphics{draft_report_v2_files/figure-latex/exp1-3factor-plot-1}

@@

Finally, the analysis showed a significant interaction effects of block by game ( F(2,100) = 6.92 , p = .002, \(\eta^{2}\) = 0.02), indicating that within-game learning differed between the games. Indeed, second half scores in RPS are significantly higher than first-half scores (t(150) = 5.59, p \textless{}.0001), while there was no significant difference between block scores for the other two games. This is indicative of the significant within game learning happening in the first game when players have no epxerience against the opponent, as opposed to much lower within game performance improvement in the latter games when participants have had some exprience playing against the opponenet and start with higher scores indicative of transfer. There was also a three-way interaction between condition, game, and block ( F(2,100) = 3.88 , p = .023, \(\eta^{2}\) = 0.01), which indicates, as seen in Figure~\ref{fig:exp1-3factor-plot} that within-game learning changes across games also depend on the sophistication of the opponent. For instance, there is more within game learning in the third game against level-2 opponents, since the initial scores are lower than agaisnt level-1 opponent. The explanation for this will become clearer when we discuss the factors moderating learning transfer in the next section.

@@

As a measure for learning transfer, we focus on participants' scores in the first 5 rounds excluding the initial round (rounds 2-6). We exclude the very first round for which the agent is programmed to play randomly as it has no data on prior rounds on which to build its response, and therefore we cannot expect the participants to exhibit any learned best responses either. A group of players with no experience of the game are expected to have scores not significantly different from 0 over the early rounds of a new game, as was the case in RPS. Any significantly positive group average scores would therefore reflect prior learning from past experiences.

@@

{\centering \includegraphics{draft_report_v2_files/figure-latex/exp1-score-by-opp-1}

@@

{\centering \includegraphics{draft_report_v2_files/figure-latex/exp2-score-by-opp-1}

@@

{\centering \includegraphics{draft_report_v2_files/figure-latex/exp2-early-scores-1}

@@

{\centering \includegraphics{draft_report_v2_files/figure-latex/exp1-comp-models-1}

@@

\caption{Experiment 1 - Histogram of best fitting computational models by condition}\label{fig:exp1-comp-models}
\end{figure}

@@

{\centering \includegraphics{draft_report_v2_files/figure-latex/exp1-cumScores-1}

@@

\caption{Experiment 1 - Average cumulative scores of participants by best fitting model}\label{fig:exp1-cumScores}
\end{figure}

However, this is at odds with the findings from the previous section regarding learning transfer: If indeed most participants use Q-learning with states to choose their actions, then they should not be able to transfer learning to the early rounds of the new game. In order to understand better what is going on, we plot the likelihood by trial for each game and each of the three strategies: Q-learning with states, and Theory of Mind models with and without the possibility of across game transfer.

Figure~\ref{fig:exp1-lik-by-tr} shows that in the later games, the likelihood for the ToM models is higher in the initial rounds in which learning transfer is measured, but that over time, the likelihood of Q-learning model becomes more important and exceeds that of ToM models.

{\centering \includegraphics{draft_report_v2_files/figure-latex/exp1-lik-by-tr-1}

\caption{Experiment 1 Likelihood by trial by game and opponent faced}\label{fig:exp1-lik-by-tr}
\end{figure}

@@

{\centering \includegraphics{draft_report_v2_files/figure-latex/exp2-comp-models-1}

@@

\caption{Experiment 2 Histogram of best fitting computational models by condition}\label{fig:exp2-comp-models}
\end{figure}

@@

{\centering \includegraphics{draft_report_v2_files/figure-latex/exp2-cumScores-1}

@@

\caption{Experiment 2 Average cumulative scores of participants by best fitting model}\label{fig:exp2-cumScores}
\end{figure}

Plotting cumulative scores by best model for experiment 2, we see very similar results looking at Figure~\ref{fig:exp2-cumScores}, in that participants whose behavior was best fit by a ToM model of learning the oppponent strategy had the highest cumulative performance. Out of these ToM models, the one in which there is within-game but no between-game transfer (NBT) had the best cumulative performance (although it only fit 2 participants best), followed by a model in which both within and between trasnfer of opponent models is allowed (BT). The next best model from a performance perspective was a Q-learning model with states and within game transfer, followed by ToM models where either players started learning from scratch at each stage of each game (NT) and the model in which they didn't distinguish between opponenets (Naive). As expected, random play was at the bottom of cumulative performance.

As in experiment 1, we want to understand the dynamic of strategy choice by plotting the likelihood by trial for each strategy, using the optimal parameters found when fitting the model. Figure~\ref{fig:exp2-lik-by-tr} shows that, as in experiment 1, ToM models had higher likelihood in the early stages of the second (most similar) game, however the likelihood of Q-learning with states models increases steadily to be the highest in the later stages of all games. In the third and mosre dissimilar game, we get a result that is different from experiment 1. In this instance, the likelihoods of the ToM models stay constant and close to their initial values.

{\centering \includegraphics{draft_report_v2_files/figure-latex/exp2-lik-by-tr-1}

\caption{Experiment2 likelihood by trial by game and opponent faced}\label{fig:exp2-lik-by-tr}
\end{figure}

\hypertarget{using-hidden-markov-model-to-explore-strategy-switching}{%
\subsection{Using Hidden Markov Model to explore strategy switching}\label{using-hidden-markov-model-to-explore-strategy-switching}}

The fact that the likelihoods of the main strategies considered cross over in both experiments could be interpreted as indicative of participants switching between strategies as the games progressed. Indeed, in both experiments, following our results, it seems that in the earlier stages of the latter games, the ToM based strategies fit observed action choices better than Q-learning based ones, with a reversal of the roles in later stages.

In order to test for the existence of strategy switching in participants' play, we fit Hidden Markov Models in which the latent states are the 3 strategies used (Q learning with state space consisting of previous round play, ToM based model with and without opponent model transfer). Hidden Markov models are useful tools to explore structure in observed time series. They are named because of two properties: First, they make the assumption that any observable action at time t results from a process whose state at time t , named \(S_{t}\) is \enquote{hidden} from the observer. Second, it also assumes that this hidden process has a Markov property, meaning that given state \(S_{t-1}\), the value of \(S_{t}\) is independent of all states occuring before time \(t-1\). We also assume that \(S_{t}\) has a discrete probability distribution in that it take one of K discrete values. The model is therefore specified by initial probabilities of being in each state \({1,2,...,K}\) and transition probabilities for moving from state \(i\) to state \(j\). These probabilities are fit using obsserved actiosn generated from these hidden states.

Using the R DepmixS4 package, we fit a total of two HMMs. In the first model, we allow for a non-nil probability of players transitioning from one state (strategy) to another. In the second model, we assume that such switching does not happen, and as such assume implicitly that when players start with a particular strategy, they continue using it throughout the game. We then compare the likleihoods of each HMM model using a likelihood ration test. In experiment 1, the HMM model with swithcing has a significantly higher likelihood than the non-swithcing one (p-value = 0.00). The switching model in experiment two has also significantly higher likelihood (p-value = 0.00). On top of indications from looking at the liklihood by trial graphs, we have therefore further evidence that participants did indeed switch their strategies as the games progressed.

### 2020-09-08 comp_model_1019.Rmd
fixed bugs in report

## Putting results together 
```{r}
load("QL_modelling.RData")
load("QL_states_modelling.RData")
load("EWA_modelling.RData")
load("EWA_self_modelling.RData")
load("MBM_modelling.RData")
load("EWA_states_modelling.RData")

>>>>

## Putting results together 
```{r}
# load("QL_modelling.RData")
# load("QL_states_modelling.RData")
# load("EWA_modelling.RData")
# load("EWA_self_modelling.RData")
# load("MBM_modelling.RData")
# load("EWA_states_modelling.RData")

### 2020-09-08 exp1_lik_by_trial.Rmd
fixed bugs in report

  exp1_lik_by_trial <- rbind(exp1_lik_by_trial,
                        data.frame(
                          "human_id" = id,
                          "condition" = dat$condition,
                          "game" = dat$game,
                          "round" = dat$round,
                          "QLS_lik" = QLS_lik,
                          "Bayes_NT_lik" = Bayes_NT_lik,
                          "Bayes_Tr_lik" = Bayes_Tr_lik

                        ))
}

@@

which(exp1_lik_by_trial$QLS_lik == 1.0 )

>>>>

  exp1_lik_by_trial <- rbind(exp1_lik_by_trial,
                        data.frame(
                          "human_id" = id,
                          "condition" = dat$condition,
                          "game" = dat$game,
                          "round" = dat$round,
                          "QLS_lik" = QLS_lik,
                          "Bayes_NT_lik" = Bayes_NT_lik,
                          "Bayes_Tr_lik" = Bayes_Tr_lik
                        ))
  #cat(as.character(Bayes_Tr_lik[c(1,51,101)]))
}

write.csv(exp1_lik_by_trial,file="exp1_lik_by_trial.csv",row.names = FALSE)

@@

summary(exp1_lik_by_trial$QLS_lik == 1.0 )
summary(exp1_lik_by_trial$QLS_lik == 0.0 )
summary(is.na(exp1_lik_by_trial$QLS_lik))

For
